---
title: "Benchmark Validation Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark Validation Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette shows a complete validation-benchmark workflow that works even
before external gold-standard datasets are attached.

## 1. Create a benchmark bundle

```{r}
library(PhysioMoCap)

ex <- createBenchmarkExample(
  n_trials = 3,
  n_samples = 250,
  noise_sd = 0.02,
  seed = 1
)

ex$manifest
```

## 2. Validate manifest integrity

```{r}
v <- validateBenchmarkManifest(ex$manifest, data_dir = ex$data_dir)
v
```

## 3. Run benchmark suite

```{r}
suite <- runBenchmarkSuite(
  manifest = ex$manifest,
  data_dir = ex$data_dir,
  thresholds = defaultBenchmarkThresholds("balanced"),
  alignment = "truncate"
)

suite
suite$suite_summary
head(suite$metrics)
```

## 4. Export reports

```{r}
report_dir <- tempfile("benchmark_report_")
runBenchmarkSuite(
  manifest = ex$manifest,
  data_dir = ex$data_dir,
  thresholds = defaultBenchmarkThresholds("balanced"),
  report_dir = report_dir
)

list.files(report_dir)
```

## 5. Use real gold-standard data later

1. Keep the same manifest columns (`benchmark_id`, `prediction_file`, `reference_file`).
2. Replace file paths with your real prediction/reference files (`.csv`, `.mot`, `.sto`, `.trc`).
3. Re-run `validateBenchmarkManifest()` and `runBenchmarkSuite()`.

This keeps validation infrastructure stable while datasets evolve.
