---
title: "Force Plate and Inverse Dynamics Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Force Plate and Inverse Dynamics Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette demonstrates a standard lower-limb kinetics workflow in `PhysioMoCap`:

1. Filter force-plate GRF signals.
2. Compute center of pressure (COP).
3. Extract loading rate and impulse.
4. Estimate joint moments and power with planar inverse dynamics.
5. Run multi-plate selection and 3D inverse dynamics.

## 1. Simulate force-plate inputs

```{r}
library(PhysioMoCap)

set.seed(1)
n <- 1000
sr <- 1000

forces <- cbind(
  Fx = rnorm(n, 0, 8),
  Fy = rnorm(n, 0, 8),
  Fz = c(rep(0, 200), abs(sin(seq(0, pi, length.out = 600))) * 900, rep(0, 200))
)

moments <- cbind(
  Mx = rnorm(n, 0, 25),
  My = rnorm(n, 0, 25),
  Mz = rnorm(n, 0, 10)
)
```

## 2. Run force-plate analysis

```{r}
out <- analyzeForcePlate(
  forces = forces,
  moments = moments,
  sampling_rate = sr,
  cutoff = 20,
  threshold = 20,
  filter_method = "moving_average"
)

out$summary
head(out$loading_rate)
head(out$impulse)
```

## 3. Build inverse-dynamics inputs

```{r}
t <- seq(0, (n - 1) / sr, length.out = n)

joints <- data.frame(
  ankle_x = rep(0.00, n),
  ankle_y = rep(0.05, n),
  toe_x = rep(0.15, n),
  toe_y = rep(0.01, n),
  knee_x = rep(0.00, n),
  knee_y = rep(0.45, n),
  hip_x = rep(0.00, n),
  hip_y = rep(0.85, n)
)

grf_2d <- data.frame(
  fx = out$filtered_forces$force_x,
  fy = out$filtered_forces$force_z,
  cop_x = if (!is.null(out$cop)) out$cop$cop_x else rep(0, n),
  cop_y = if (!is.null(out$cop)) out$cop$cop_y else rep(0, n)
)

angles <- data.frame(
  ankle = 0.20 * sin(2 * pi * 1 * t),
  knee = 0.45 * sin(2 * pi * 1 * t + 0.2),
  hip = 0.35 * sin(2 * pi * 1 * t + 0.4)
)
```

## 4. Estimate inertial parameters and compute moments/power

`inverseDynamics2D()` runs a recursive link-segment Newton-Euler chain
(foot -> shank -> thigh), so it needs each segment's mass and centre of mass as
well as the foot distal end (`toe_x`, `toe_y`). The legacy massless-segment
approximation is still reachable as `model = "quasi_static"`, but it omits
segment weight and inertia entirely - it returns zero moments through the whole
swing phase - and is kept only for reproducing older analyses.

```{r}
inertial <- estimateSegmentInertia(
  body_mass = 70,
  segment_lengths = c(foot = 0.25, shank = 0.43, thigh = 0.45)
)

id_out <- inverseDynamics2D(
  joints = joints,
  grf = grf_2d,
  sampling_rate = sr,
  angles = angles,
  inertial = inertial
)

# The moving-average filter leaves NA at the ends of the force trace, so look
# at frames from the middle of the trial.
id_out[seq(round(n / 2), length.out = 6), ]
```

The moments jitter from frame to frame because the simulated plate moments are
white noise, so the centre of pressure jitters too and swings the ground
reaction force about the joint centres. Real data needs the forces, the centre
of pressure and the markers low-pass filtered before inverse dynamics.

## 5. Multi-plate selection and 3D inverse dynamics

```{r}
force_z <- cbind(
  fp1 = out$filtered_forces$force_z * 0.6,
  fp2 = out$filtered_forces$force_z
)

pe_fp <- PhysioCore::PhysioExperiment(
  assays = S4Vectors::SimpleList(
    force_x = cbind(fp1 = out$filtered_forces$force_x * 0.6,
                    fp2 = out$filtered_forces$force_x),
    force_y = cbind(fp1 = out$filtered_forces$force_y * 0.6,
                    fp2 = out$filtered_forces$force_y),
    force_z = force_z
  ),
  colData = S4Vectors::DataFrame(
    label = c("fp1", "fp2"),
    type = c("forceplate", "forceplate")
  ),
  samplingRate = sr
)

fp_auto <- analyzeForcePlatePE(pe_fp, plate_index = "auto", threshold = 20,
                               cutoff = 20, filter_method = "moving_average")
fp_auto$selected_plate
```

```{r}
# Keep the vertical along y so that it matches `vertical = "y"` below, and add
# a lateral (z) coordinate.
joints_3d <- transform(
  joints,
  ankle_z = 0, toe_z = 0, knee_z = 0, hip_z = 0
)

grf_3d <- data.frame(
  fx = out$filtered_forces$force_x,
  fy = out$filtered_forces$force_z,
  fz = out$filtered_forces$force_y,
  cop_x = rep(0, n),
  cop_y = rep(0, n),
  cop_z = rep(0, n)
)

angles_3d <- data.frame(
  ankle_x = angles$ankle, ankle_y = 0 * angles$ankle, ankle_z = 0 * angles$ankle,
  knee_x = angles$knee,   knee_y = 0 * angles$knee,   knee_z = 0 * angles$knee,
  hip_x = angles$hip,     hip_y = 0 * angles$hip,     hip_z = 0 * angles$hip
)

id3 <- inverseDynamics3D(
  joints = joints_3d,
  grf = grf_3d,
  sampling_rate = sr,
  angles = angles_3d,
  inertial = inertial,
  vertical = "y"
)

id3[seq(round(n / 2), length.out = 6), 1:7]
```

## Notes

- For publication-grade kinetics, verify sign conventions and coordinate frames
  from your lab's force-plate and marker setup.
- Use subject-specific segment lengths and MVC/body-mass normalization where
  appropriate.
