---
title: "Getting Started with PhysioMoCap"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with PhysioMoCap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This guide is designed for first-time users.

## 1. Create demo data

```{r}
library(PhysioMoCap)

demo <- demoMoCapData(seed = 1)
class(demo$mocap)
head(demo$joints)
```

## 2. Run one-command quick start

```{r}
qs <- quickStartMoCap(seed = 1)
qs
```

## 3. Check data readiness

```{r}
qs$readiness
head(qs$readiness$checks)
```

## 4. Inspect key outputs

```{r}
qs$forceplate$summary
head(qs$inverse_dynamics)
```

## 5. Start from your own file

```{r}
trc_file <- system.file("testdata", "sample.trc", package = "PhysioMoCap")
if (nzchar(trc_file)) {
  pe <- readMoCapAuto(trc_file)
  assessMoCapReadiness(pe)
}
```

You can also run `quickStartMoCap(path = "trial.c3d")` directly.

## 6. Common input entry points

- Marker trajectories: `readMoCapAuto()`, `readC3D()`, `readMoCapCSV()`, `readTRC()`
- OpenSim kinematics/forces: `readMOT()`, `readSTO()`
- OpenCap cloud data: `readOpenCap()`

## 7. Common troubleshooting

- If `sampling_rate` is missing, set it explicitly.
- If `detectEvents()` cannot find signals from a matrix, pass a named `signals` list.
- If optional dependency packages are missing (e.g., `c3dr`, `signal`), install them from CRAN.
